Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design

### 2025-04-04 src/main.tex
[text] table 1

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate} \todo{more}.
To tackle this issue, the conference \adae\ committees have been advising authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' made of a list of commands to execute from a ``base image'' (for example, \df\ for Docker containers).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
These base images can also be removed by the administrators making all the recipes unbuildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus threatens the reproducibility of the results.
One solution would be for the authors to store the built container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
But this come with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

The workflow of \ecg\ is depicted in Figure~\ref{fig:workflow}.
% We make sure to use an empty local Docker cache before building the containers.
The workflow itself is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment is managed with Nix~\cite{dolstra_nix_2004}.

@@

We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build the associated \df s once every month for seven months.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
%\todo{when was the artifact submission, and publicaiton of the papers for europar?}
The data collected is available on Zenodo~\cite{dataset_europar24}. \todo{update zenodo with the april measures}

@@

\begin{table}
  \centering
  \scalebox{0.850}{
  \begin{tabular}{llll}
    \toprule
    Artifact & Base image & Version & \texttt{update}?\\
    \midrule
    \texttt{canon\_solving}~\cite{canon_solving}   & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \texttt{geijer\_how}~\cite{geijer_how}         & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \texttt{hiraga\_peanuts}~\cite{hiraga_peanuts} & \texttt{devcontainers/cpp} & \texttt{1-debian-12} & Yes \\
    \texttt{munoz\_fault}~\cite{munoz_fault}       & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \texttt{wolff\_fast}~\cite{wolff_fast}         & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \bottomrule
  \end{tabular}
  }
  \caption{Information about the \df s from the study.}\label{tab:info}
\end{table}

@@

\todo{update with april measures}
The packages that changed the most in terms of versions are:
\begin{enumerate}
\item \texttt{linux-libc-dev:amd64} (installed via \texttt{dpkg}) with 6 different versions for each artifact (1 per new build attempt)
\item \texttt{fonttools} (installed via \texttt{pip}) with 5 different versions for the artifacts \texttt{geijer\_how} and \texttt{wolff\_fast}
\item \texttt{numpy} (installed via \texttt{pip}) with 5 different versions for the artifact \texttt{wolff\_fast}
\end{enumerate}

>>>>

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment \cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To tackle this issue, the conference \adae\ committees have been advising authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' made of a list of commands to execute from a ``base image'' (for example, \df\ for Docker containers).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
These base images can also be removed by the administrators making all the recipes unbuildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus threatens the reproducibility of the results.
One solution would be for the authors to store the built container itself (\ie\ the tarball) onto some long term storage for example Zenodo~\cite{zenodo}.
But this come with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

The workflow of \ecg\ is depicted in Figure~\ref{fig:workflow}.
% We make sure to use an empty local Docker cache before building the containers.
The workflow itself \cite{ecg_workflow} is ruled by a Snakemake~\cite{koster2012snakemake} description and its software environment \cite{ecg_nix} is managed with Nix~\cite{dolstra_nix_2004}.

@@

We wrote the descriptions in Nickel (see Section~\ref{sec:nickel}) for the five artifacts of the EuroPar'24 conference that used a Docker container to share their software environments, and used \ecg\ to build the associated \df s once every month for seven months.
% "2024-10-01" "2024-11-04" "2024-12-02" "2025-01-02" "2025-02-05" "2025-03-12"
The initial build was October 1$^{st}$ 2024, then the following builds were on November 4$^{th}$ 2024, December 2$^{nd}$ 2024, January 2$^{nd}$ 2025, February 5$^{th}$ 2025, March 12$^{th}$ 2025, and April 1$^{st}$ 2025.
The deadline to submit the artifacts was May 15$^{th}$ 2024\footnote{\url{https://2024.euro-par.org/calls/artifacts/}}.
%\todo{when was the artifact submission, and publicaiton of the papers for europar?}
The data collected is available on Zenodo~\cite{dataset_europar24}.

@@

\begin{table}
  \centering
  \scalebox{0.800}{
  \begin{tabular}{llll}
    \toprule
	\multirow{2}{*}{Artifact} & \multicolumn{2}{c}{Docker Base Image used} & \multirow{2}{*}{
					\begin{tabular}{c}Calling\\ \texttt{apt update}? \end{tabular}
}\\
\cmidrule(lr){2-3}
     & Name & Version & \\
    \midrule
    \texttt{canon\_solving}~\cite{canon_solving}   & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \texttt{geijer\_how}~\cite{geijer_how}         & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \texttt{hiraga\_peanuts}~\cite{hiraga_peanuts} & \texttt{devcontainers/cpp} & \texttt{1-debian-12} & Yes \\
    \texttt{munoz\_fault}~\cite{munoz_fault}       & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \texttt{wolff\_fast}~\cite{wolff_fast}         & \texttt{ubuntu} & \texttt{22.04} & Yes \\
    \bottomrule
  \end{tabular}
  }
  \caption{Information about the \df s from the study.}\label{tab:info}
\end{table}

@@

The packages that changed the most in terms of versions are:
\begin{enumerate}
\item \texttt{linux-libc-dev:amd64} (installed via \texttt{dpkg}) with 7 different versions for each artifact (1 per new build attempt), except for \texttt{hiraga\_peanuts} where there were 4 different versions.
\item \texttt{numpy} (installed via \texttt{pip}) with 6 different versions for the artifact \texttt{wolff\_fast}
\item \texttt{fonttools} (installed via \texttt{pip}) with 5 different versions for the artifacts \texttt{geijer\_how} and \texttt{wolff\_fast}
\end{enumerate}
